Live and Let Die? (Delayed Issues not Harder to Resolve)

### 2015-08-02 data.tex
Updates from ShareLaTeX

Before presenting our analysis of phase delay in software projects, we define our terms, provide an overview of TSP and the sample projects, and describe the data collection process.

\subsection{Definitions}
A \emph{defect} is any change to a product, after its construction, that is necessary to make the product correct.  A typographical error found in review is a defect. If that same defect is discovered while writing the code but before review, it is not considered to be a defect.

The \emph{cost-to-fix} can be computed in two ways. The first is a direct measure of the actual work time spent in correcting the defect in the product. Direct measurement enables analysis of the variability the cost-to-fix among a category of defects. The second approach measures the total time spent in a defect removal phase and divides by the total defects removed. This provides an average cost-to-fix including overhead of that activity. For example, using method two, a code inspection may require an hour of effort to remove three defects for a 3 defects per hour or to invert the ratio, 1/3 hours per defect. Most of the time, however, is spent locating the defects not actually fixing them. we use the second approach in our analysis.

\subsection{Overview of the Team Software Process}
To empirically test for the phase delay effect, we examined 171 software projects conducted between 2005 and 2014. These projects took place at organizations in many countries and were conducted using  the Team Software Process (TSP$\textsuperscript{SM}$), a software project management approach developed at the Software Engineering Institute (SEI) at Carnegie Mellon University~\cite{tsp00}. TSP is an extension of the Personal Software Process (PSP$\textsuperscript{SM}$) developed at the SEI by Watts Humphrey ~\cite{tsp00}. The data from these TSP projects were collected and stored in the Software Engineering Measured Process Repository (SEMPR) at the SEI. The Software Engineering Institute (SEI) at Carnegie Mellon University explores methods for software process improvement.

TSP has guiding principles; the following are relevant to this study:
\begin{itemize}
\item  Sound engineering work does not happen by accident. It must be planned.
\item  Performance of the plans requires realistic commitments of the resources needed to perform the work.
\item  The planned work should frequently be compared to  results to assure plans remain relevant.
\item  People performing the work should be responsible for planning and tracking the work.
\end{itemize}

TSP helps developers by providing a set of measures that can be applied to managing tasks, quality, and schedule. Planning begins by quantifying goals, defining work practices, and estimating size and effort. Developers then use this information to make a detailed short term plan. As the developers perform project work, they use a tool such as the Process Dashboard to collect their time, effort, size, and schedule data. Every week, the team reviews their data to evaluate status, identify actual rates, determine if project goals for schedule, cost, and quality are being met. The team then uses this information to make necessary plan corrections. At the end of the project the coach and team perform a quantitative project post mortem.

@@

{\em Planning} follows the technique taught in the PSP$\textsuperscript{SM}$. Developers estimate the size of work products and convert this to a total effort using historical rates. Time in specific tasks come from the  process phases and historical percent time in phase distributions. Defects are estimated using historical phase injection rates and phase removal yields. Coaches help the developers to compare estimates against actual results. In this way, developers acquire a more realistic understanding of their work behavior, performance, and schedule status.

@@

\begin{figure}[!t]  
\begin{center}
\includegraphics[width=2.5in]{img/waterfall-v3.png}  
\end{center}
\caption{Phases of our data.
Abbreviations: \newline
{\em Before}= before development;\newline
{\em Reqts}	  = requirements;\newline
{\em HLD}	  = high-level design;\newline
{\em IntTest} = Integration testing (with code from others);\newline
{\em SysTest} = system test (e.g. load stress tests);\newline
{\em AcceptTest}  = acceptance testing (with users);\newline
{\em review}        = private activity;\newline
{\em inspect}        = group activity.}
\label{fig:waterfall}
\end{figure}

\subsection{Projects in the Sample}
Since shortly before the year 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has been been mentoring software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2014.
The projects were mostly small to medium, with a median duration of 46 days and a maximum duration of 90 days in major increments. 
Several projects extended for multiple incremental development cycles. 
Median team size was 7 people, with a maximum of 40.

Many of the projects were e-commerce web portals or banking systems in the US, South Africa, and Mexico. 
There were  some  medical device projects in  the US, France, Japan, and Germany as well  as a commercial computer-aided design systems, and embedded systems.

An anonymized version of that data is available in the PROMISE repository\footnote{http://openscience.us/repo}  and,
for confidentiality restrictions, we cannot offer 
further details on these projects.

\subsection{Data Collection Process}
\label{sec:data-collection}

Organizations using TSP agree to provide their project data to the SEI for use in research. In return the SEI agrees    that  data must not be traceable to its source. The data is collected at major project events; launch, interim checkpoints, and at project completion. Data includes project and site  characteristic summaries, surveys of team members, launch presentations, launch outbriefs, and baseline plans, final data from the project, and the project post mortem report.  In practice, this data requirement has only been enforced for the purposes of certifying and reauthorizing TSP coaches who must submit data to maintain their authorization. Coaches are certified by demonstrating competent use of the TSP process with the artifacts and data not by the actual project results.  Of the data submitted, only the data recorded using the Process Dashboard tool has been collected and aggregated for this research.

To use this data for research and benchmarking,  Nichols first collected the Process Dashboard data from the submission repository then used a tool built by the Process Dashboard developer, David Tuma, 
to gather the project data into a database. Nichols and his colleague Yasutaka Shirai then extracted data into views suitable for analysis. The key views include a project results summary, effort logs, task logs, and defect logs. The data has not yet undergone additional screening. A summary of the data quality issues was reported in a previous work ~\cite{shirai14} (that summary is discussed further in our {\em Validity} section, below).

The tool collected process information about those tasks and defects that we analyze in this work. 
That task data include  work start time, work end time, delta
work time, and interruption time. Software engineers are often
interrupted by meetings, requests for technical help, reporting, and
so forth. These events are recorded, in minutes, as interruption
time.

 Defect logs include the time and date a defect was discovered, the phase in which that defect was injected, the phase in which it was removed, the time (in minutes) required to find and fix the defect, and the categorical type.

The defect types in the most common defect type standard  in the the SEI TSP data are follows:
\begin{itemize}
\item Environment: design, compile, test, or other support system problems
\item Interface: procedure calls and reference, I/O, user format
\item Data: structure, content
\item Documentation: comments, messages
\item Syntax: spelling, punctuation typos, instruction formats
\item Function: logic, pointers, loops, recursion, computation, function defects  
\item Checking: error messages, inadequate checks
\item Build: change management, library, version control
\item Assignment: package
declaration, duplicate names, scope, limits
\item System: configuration, timing, memory
\end{itemize}

As of November 2014, the SEI TSP database contained data from 212
TSP projects. The projects completed between July 2006 and
November 2014; they included 47 organizations and 843 people. 
The database fact tables
contain 268,726 time logs, 
154,238 task logs,
 47,376 defect logs, 
and 26,534 size logs. 
After selecting defects from the data log and joining the data to the time log table  171 of these projects remained. The excluded projects had no or too few defects to use in this analysis.

In this paper, when we report ``time to resolve an
issue,'' we show the difference between the start and end times
of a work session, with any interruption time subtracted (the
difference in times, minus the interruptions).

%That said, certain semantic features of the SEI TSP data should be noted.
%Firstly, in the current TSP collection tool, 
%fix times are only the developer time for the developer walking through the phases of \fig{waterfall}.
%We are currently tracking the fix time for post-release issues (e.g. those raised during  acceptance test and %later
%product life cycle). So far, in that post-release data,  we have not detected
%a dramatic phase escalation effect (but at this time, we have nothing definitive comment on that matter).%
%
%\bill{somewhere you have one note on \underline{find} and fix times.  for this paper, we need just fix times. %but is there
%anything we need to fret about re \underline{find} times?}

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The logical phases used in this paper are shown in \fig{waterfall}. Although the representation suggests a waterfall, the phases represent the primary stages through which requirements are developed into working and tested code. That is, all requirements must be stated in some way, implemented in code, integrated, and tested. All real implementations of any size follow a spiral approach with many team performing the work in iterative and/or incremental development cycles.  Within a project increment, multiple features or components may be developed and incremented. TSP is compatible with agile and encourages iterative and incremental development. Nonetheless, the specific strategy and cycle duration is a project decision. TSP does, however, strongly encourage 1) constructing units of sufficient size that measurement is practicable, and 2) separating the construction from appraisal and test activities. This effectively highlights the separation of construction from rework activities and aids the apportionment of defects to those found in appraisal activities (reviews and inspections) and those found through failure (test). The distinct construction activities (requirements, high and detailed design, and code) were chosen to help teams analyze the effectiveness and efficiency of their practices through analysis of the defect phase origin, type, fix effort, and phase of discovery.

Note that, in that Figure~\ref{fig:waterfall}:
\bi 
\item
Several  phases in which product is created have sub-phases of {\em review} and {\em inspect} to remove defects. TSP uses review, a sub-phase in which individuals perform personal reviews of their work products prior to the peer review, which TSP calls the inspection.
\item Testing is divided into several stages. Developers perform unit test prior to code complete.  After code complete a standard phase is the integration, which combines program units into workable system ready for system test. Integration,  system test, and acceptance test are often performed by a separate group. 
\ei
For these TSP projects, the principles followed are: 
\bi 
%\item Defects should be found before test, that is, test defects should be infrequent.
\item A product should be inspected before used in a subsequent construction phase to minimize  unnecessary rework.
\item To minimize unnecessary rework, a product should be reviewed and inspected prior to use in a subsequent construction phase.
\item A personal review should remove the most common defects before being given to peers for the inspection.
\item Because different categories of test will find different types of defect, and follow in a natural sequence, the tests should be measured separately
\ei 
In summary, the TSP process and measurement have been specifically designed to aid the analysis of defects for use in process improvement. The data are, therefore, uniquely suitable for an analysis of the phase delay effect. 
\subsection{Distinguishing Project Characteristics}
Modern development processes include the core engineering practices of estimation, planning, configuration control, requirements, design, code development, and test. Agile methods address  not only these core practices but also address team organization, individual and team commitments, test and build automation, and implementing iterative and incremental development. TSP uses additional practices associated with high performance ~\cite{jones10} including; project tracking and control, specialization of team members, inspections, static analysis, documentation and training, and reuse. Although the teams in this sample use many typical agile practices, they are also using additional practices.

>>>>

The above analysis motivates a more detailed look at the delayed issued effect.  
Accordingly, we examined 171 software projects conducted between 2006 and 2014. These projects took place at organizations in many countries and were conducted using  the Team Software Process (TSP$\textsuperscript{SM}$).

Since 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has been been mentoring software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2014.
The projects were mostly small to medium, with a median duration of 46 days and a maximum duration of 90 days in major increments. 
Several projects extended for multiple incremental development cycles. 
Median team size was 7 people, with a maximum of 40. 
Many of the projects were e-commerce web portals or banking systems in the US, South Africa, and Mexico. 
There were  some  medical device projects in  the US, France, Japan, and Germany as well  as a commercial computer-aided design systems, and embedded systems.

An anonymized version of that data is available in the PROMISE repository at openscience.us/repo.
For confidentiality restrictions, we cannot offer 
further details on these projects.

\subsection{About TSP$\textsuperscript{SM}$}

TSP is a software project management approach developed at the Software Engineering Institute (SEI) at Carnegie Mellon University~\cite{tsp00}. TSP is an extension of the Personal Software Process (PSP$\textsuperscript{SM}$) developed at the SEI by Watts Humphrey ~\cite{tsp00}. The data from these TSP projects were collected and stored in the Software Engineering Measured Process Repository (SEMPR) at the SEI. The Software Engineering Institute (SEI) at Carnegie Mellon University explores methods for software process improvement.

%TSP helps developers via a set of measures that can be applied to managing tasks, quality, and schedule. Planning begins by quantifying goals, defining work practices, and estimating size and effort. Developers then use this information to make a detailed short term plan. As the developers perform project work, they use a tool such as the Process Dashboard to collect their time, effort, size, and schedule data. Every week, the team reviews their data to evaluate status, identify actual rates, determine if project goals for schedule, cost, and quality are being met. The team then uses this information to make necessary plan corrections. At the end of the project the coach and team perform a quantitative project post mortem.

@@

During {\em Planning}, developers estimate the size of work products and convert this to a total effort using historical rates. Time in specific tasks come from the  process phases and historical percent time in phase distributions. Defects are estimated using historical phase injection rates and phase removal yields. Coaches help the developers to compare estimates against actual results. In this way, developers acquire a more realistic understanding of their work behavior, performance, and schedule status.

@@

\subsection{Data Integrity}

A common property of real-world data sets is the presence
of noisy entries (superfluous  or spurious data). 
The level of noise can be quite high. As reported
in \cite{shepperd12}, around
10\% to 30\%
of the records in the NASA MDP defect data sets are
affected by noise.

One reason to use the SEI data for the analysis of this paper is its remarkably low level of noise.
Nichols et al.~\cite{shirai14}  report that
the noise levels in the SEI TSP data are smaller than those seen
in other data sets. They found in the SEI TSP data that:\bi 
\item
4\% of the data was incorrect (e.g. nulls, illegal formats);
\item  2\% of the data has inconsistencies such as timestamps
where the stop time was before the start time;
\item 3\% of the data had data that was not credible
such as tasks listed in one day that took like than six hours.
\ei 
One explanation for this low level of noise is the TSP process.
One the guiding principles of TSP was that  people performing the work are  responsible for planning and tracking the work. That is,  all the data collected here was entered
by local developers. This data was then check by local {\em coaches} before being sent to the SEI
databases. Coaches are certified by demonstrating competent use of the TSP process with the artifacts and data.
The use of certified local coaches within each project increases the integrity of our data.

Overall, the   effort associated with adding TSP to a project is not onerous. McHale reports~\cite{mchale02}:
\bi
\item
 The time spent  tracking time, defects, and tasks requires less then 3\% of a developer's time. Weekly team meetings  require at most an hour, which is
only 2.5\% of a 40 hour work week. 
\item
Team launches and replans average about 1 day per month or 5\% planning overhead.
\ei
It is true that one staff member is needed as a ``couch'' to mentor the teams
and certify and monitor that data collection. However, one of us (Nichols) has worked with dozens of TSP teams. He reports that one  trained coach can support 4 or 6 teams (depending upon team experience).

\subsection{Data Details}
\label{sec:data-collection}

%Organizations using TSP agree to provide their project data to the SEI for use in research. In return the SEI agrees    that  data must not be traceable to its source. The data is collected at major project events; launch, interim checkpoints, and at project completion. Data includes project and site  characteristic summaries, surveys of team members, launch presentations, launch outbriefs, and baseline plans, final data from the project, and the project post mortem report.  In practice, this data requirement has only been enforced for the purposes of certifying and reauthorizing TSP coaches who must submit data to maintain their authorization. Coaches are certified by demonstrating competent use of the TSP process with the artifacts and data not by the actual project results.  Of the data submitted, only the data recorded using the Process Dashboard tool has been collected and aggregated for this research. 
 %The key views include a project results summary, effort logs, task logs, and defect logs. The data has not yet undergone additional screening. A summary of the data quality issues was reported in a previous work ~\cite{shirai14} (that summary is discussed further in our {\em Validity} section, below).

Using tools provided by the SEI, developers kept very detailed logs of their daily activity. Our data includes  work start time, work end time,  delta
work time, and interruption time. Software engineers are often
interrupted by meetings, requests for technical help, reporting, and
so forth. These events are recorded, in minutes, as interruption
time.  
In this paper, when we report ``time to resolve an
issue,'' we show the difference between the start and end times
of a work session, with any interruption time subtracted (the
difference in times, minus the interruptions).

As of November 2014, the SEI TSP database contained data from 212
TSP projects. The projects completed between July 2006 and
November 2014; they included 47 organizations and 843 people. 
The database fact tables
contain 268,726 time logs, 
154,238 task logs,
 47,376 defect logs, 
and 26,534 size logs. 
After selecting defects from the data log and joining the data to the time log table  171 of these projects remained (the excluded projects had no or too few defects to use in this analysis).

\begin{figure}[!b]  
\begin{center}
\includegraphics[width=2.5in]{img/waterfall-v3.png}  
\end{center}
\caption{Phases of our data.
Abbreviations: \newline
{\em Before}= before development;\newline
{\em Reqts}	  = requirements;\newline
{\em HLD}	  = high-level design;\newline
{\em IntTest} = Integration testing (with code from others);\newline
{\em SysTest} = system test (e.g. load stress tests);\newline
{\em AcceptTest}  = acceptance testing (with users);\newline
{\em review}        = private activity;\newline
{\em inspect}        = group activity.}
\label{fig:waterfall}
\end{figure}

In these logs, a  \emph{defect} is any change to a product, after its construction, that is necessary to make the product correct.  A typographical error found in review is a defect. If that same defect is discovered while writing the code but before review, it is not considered to be a defect. 
SEI TSP defect types are:
\begin{itemize}
\item Environment: design, compile, test,  other support  problems
\item Interface: procedure calls and reference, I/O, user format
\item Data: structure, content
\item Documentation: comments, messages
\item Syntax: spelling, punctuation typos, instruction formats
\item Function: logic, pointers, loops, recursion, computation  
\item Checking: error messages, inadequate checks
\item Build: change management, library, version control
\item Assignment: package
declaration, duplicate names, scope
\item System: configuration, timing, memory
\end{itemize}
The defect logs in that data  include the time and date a defect was discovered, the phase in which that defect was injected, the phase in which it was removed, the time (in minutes) required to find and fix the defect, and the categorical type.

The phases logged by our data are shown in \fig{waterfall}.
 Although the representation suggests a waterfall model, the SEI experience is that all real implementations of any size follow a spiral approach with many team performing the work in iterative and/or incremental development cycles. %Within a project increment, multiple features or components may be developed and incremented. TSP is compatible %with agile and encourages iterative and incremental development. Nonetheless, the specific strategy and cycle %duration is a project decision. TSP does, however, strongly encourage 1) constructing units of sufficient size that %%measurement is practicable, and 2) separating the construction from appraisal and test activities. This %effectively highlights the separation of construction from rework activities and aids the apportionment of defects %to those found in appraisal activities (reviews and inspections) and those found through failure (test). The %distinct construction activities (requirements, high and detailed design, and code) were chosen to help teams %analyze the effectiveness and efficiency of their practices through analysis of the defect phase origin, type, fix %effort, and phase of discovery.

Note that, in that Figure~\ref{fig:waterfall},
several  phases in which product is created have sub-phases of {\em review} and {\em inspect} to remove defects.
In TSP  reviews,  individuals perform personal reviews of their work products prior to the peer review
(which TSP calls the inspection(.
Also, Figure~\ref{fig:waterfall} divides  testing     as follows. Developers perform unit test prior to code complete.  After code complete a standard phase is the integration, which combines program units into workable system ready for system test. Integration,  system test, and acceptance test are often performed by another group.

%For these TSP projects, the principles followed are: 
%%\bi 
%\item Defects should be found before test, that is, test defects should be infrequent.
%%\item A product should be inspected before used in a subsequent construction phase to minimize  unnecessary rework.
%\item To minimize unnecessary rework, a product should be reviewed and inspected prior to use in a subsequent construction phase.
%\item A personal review should remove the most common defects before being given to peers for the inspection.
%\item Because different categories of test will find different types of defect, and follow in a natural sequence, the tests should be measured separately
%\ei 
%In summary, the TSP process and measurement have been specifically designed to aid the analysis of defects for use in process improvement. The data are, therefore, uniquely suitable for an analysis of the phase delay effect. 
%\subsection{Distinguishing Project Characteristics}
%Modern development processes include the core engineering practices of estimation, planning, configuration control, requirements, design, code development, and test. Agile methods address  not only these core practices but also address team organization, individual and team commitments, test and build automation, and implementing iterative and incremental development. TSP uses additional practices associated with high performance ~\cite{jones10} including; project tracking and control, specialization of team members, inspections, static analysis, documentation and training, and reuse. Although the teams in this sample use many typical agile practices, they are also using additional practices.

### 2015-08-02 lucas.tex
Updates from ShareLaTeX

Our survey collected data on software engineers' views of commonly held software engineering ``laws''.
One of the laws questioned relates directly to the problems associated with
the delayed issue effect: ``Requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' (from Glass~\cite{glass02} p.71 who references Boehm \& Basili~\cite{boehm01}). We abbreviate this law as RqtsErr
(and is
illustrated in \fig{b81}).

@@

\begin{tabular}{l|c|c|c|c|c}
 &  & \multicolumn{2}{c}{agreement} & \multicolumn{2}{c}{applicability} \\\cline{1-1} 
Practitioner survey  & N & $\bar{x}$ & $\tilde{x}$ & $\bar{x}$ & $\tilde{x}$ \\
\hline 
\textbf{Rqts errors are most expensive...} & 16 & 4.6 & 5 & 4.2 & 4 \\ 
Process maturity improves output & 17 & 4.4 & 4 & 3.9 & 4 \\ 
Most time is spent removing errors & 16 & 4.1 & 4 & 3.9 & 4 \\ 
Missing reqts are hardest to fix & 17 & 4.1 & 4 & 4.0 & 4 \\
Reuse increases prod. and qual. & 16 & 3.9 & 4 & 3.8 & 4 \\
OO-programming reduces errors & 13 & 3.9 & 4 & 3.5 & 4 \\
80-20 rule (defects to modules) & 12 & 3.8 & 4 & 4.0 & 4 \\
Adding manpower to a late project & 15 & 3.7 & 4 & 3.7 & 4 \\
Inspections can remove 90\% of defects & 18 & 3.3 & 4 & 3.5 & 4 \\
Smaller changes have higher error density & 14 & 2.8 & 3 & 3.4 & 3.5 \\
A developer is unsuited to test own code & 17 & 2.6 & 3 & 3.5 & 4\\\hline

@@

Overall, the RqtsErr law was the most agreed upon and most applicable law of 11 surveyed amongst practitioners, and the second most agreed upon law amongst researchers. The survey results strongly support that the notion that the  delayed issue effect   is a widely-held belief in software engineering.

>>>>

Our survey collected data on software engineers' views of commonly held software engineering ``laws''.
One of the laws questioned is the a specific form of our delayed issue effect:  ``requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' (from Glass~\cite{glass02} p.71 who references Boehm \& Basili~\cite{boehm01}). We abbreviate this law as RqtsErr
(and is
illustrated in \fig{b81}).

(Technical aside: we    ask them  RqtsErr  since this issue typically needs no supportive explanatory
text. On the other hand, if we had asked them about our more general term ``delayed issue
effect'', we would have had to burden our respondents with extra explanations).

@@

\begin{tabular}{l|c|c|c|c|c}
 &  & \multicolumn{2}{c|}{agreement} & \multicolumn{2}{c}{applicability} \\\cline{1-1} 
Practitioner survey  & N & w & $\bar{x}$ & $\tilde{x}$ \\
\hline 
\textbf{Rqts errors are most expensive...} & 16 & 4.6 & 5 & 4.2 & 4 \\ 
Process maturity improves output & 17 & 4.4 & 4 & 3.9 & 4 \\ 
Most time is spent removing errors & 16 & 4.1 & 4 & 3.9 & 4 \\ 
Missing reqts are hardest to fix & 17 & 4.1 & 4 & 4.0 & 4 \\
Reuse increases prod. and qual. & 16 & 3.9 & 4 & 3.8 & 4 \\
OO-programming reduces errors & 13 & 3.9 & 4 & 3.5 & 4 \\
80-20 rule (defects to modules) & 12 & 3.8 & 4 & 4.0 & 4 \\
Adding manpower to a late project & 15 & 3.7 & 4 & 3.7 & 4 \\
Inspections can remove 90\% of defects & 18 & 3.3 & 4 & 3.5 & 4 \\
Smaller changes have higher error density & 14 & 2.8 & 3 & 3.4 & 3.5 \\
A developer is unsuited to test own code & 17 & 2.6 & 3 & 3.5 & 4\\\hline

@@

Overall, the RqtsErr law was the most agreed upon and most applicable law of 11 surveyed amongst practitioners, and the second most agreed upon law amongst researchers. 
This results strongly support that the notion that the  delayed issue effect   is a widely-held belief in software engineering.

### 2015-08-02 lucasRelatedWork.tex
Updates from ShareLaTeX

Yet when we look at the literature, the evidence for
delayed issue effect is both very sparse and very old.
The first data on the difficulties of resolving delayed issues as a function of lifecycle phase date back to large systems in the late 70s from IBM~\cite{Fagan76}, TRW~\cite{Boehm76}, GTE~\cite{Daly77}, and Bell Labs~\cite{Stephenson76} (Figure~\ref{fig:cost-to-fix}). These studies suggest that the difficulty (in terms of effort) to find and fix an error monotonically increases with lifecycle phase.

@@

In the 40 years since these initial studies, few studies have been published on the difficulty to resolve issues
as a function of lifecycle phase. Boehm~\cite{Boehm80} provides data suggesting that the cost-to-fix curve for small projects (from two student projects of 2000 deliverable source instructions) is flatter than for large projects (the dashed line of Figure~\ref{fig:cost-to-fix}). But note that even in this figure
that even for the smaller projects, there is a four-fold increase (from 5 to 20) in the difficulty
associated with handling issues introduced earlier in the lifecycle.

Shull et al.~\cite{Shull02} conducted a literature survey and held a series of e-workshops with industry experts on fighting defects. Workshop participants from Toshiba and IBM reported cost-to-fix ratios between early lifecycle and post-delivery defects of 1:137 and 1:117 for large projects respectively~\cite{Shull02} -- but we note here that the raw data points were not provided (which makes confirming those numbers 
a difficult task).

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example this, \fig{steck} shows one 2004 survey that reports an exponential delayed issue  effect for 
requirements issues in eight case studies. The right-hand-side of that column shows which of those referenced studies
we can still access-- and a ``?'' denotes a references from a broken url link (note that most
of those links are broken). Note that nearly all of that data is now no longer accessible.

@@

\begin{figure}
 \includegraphics[width=3in]{img/clutterbuck.png} 
 \caption{Cost of change from an agile case study. From~\cite{Clutterbuck09}.}\label{fig:clutterbuck}
 \end{figure}

>>>>

Yet when we look at the literature, the evidence for
delayed issue effect is both very sparse and very old.
The first data on the difficulties of resolving delayed issues as a function of lifecycle phase date back to large systems in the late 70s from IBM~\cite{Fagan76}, TRW~\cite{Boehm76}, GTE~\cite{Daly77}, and Bell Labs~\cite{Stephenson76} (Figure~\ref{fig:cost-to-fix}). These studies suggest that the difficulty (in terms of effort) to find and fix an error monotonically increases with lifecycle phase.  In 1990, Boehm~\cite{Boehm80} provides data suggesting that the cost-to-fix curve for small projects (from two student projects of 2000 deliverable source instructions) is flatter than for large projects (the dashed line of Figure~\ref{fig:cost-to-fix}).

@@

In the 40 years since these initial studies, few studies have explored the difficulty to resolve issues
as a function of lifecycle phase. 
Shull et al.~\cite{Shull02} conducted a literature survey and held a series of e-workshops with industry experts on fighting defects. Workshop participants from Toshiba and IBM reported cost-to-fix ratios between early lifecycle and post-delivery defects of 1:137 and 1:117 for large projects respectively~\cite{Shull02} -- but we note here that the raw data points were not provided (which makes confirming those numbers 
a difficult task).

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example of this, \fig{steck} shows one 2004 survey that reports an exponential delayed issue  effect for 
requirements issues in eight case studies. The right-hand-side of that column shows which of those referenced studies
we can still access-- and a ``?'' denotes a references from a broken url link (note that most
of those links are broken). Note that nearly all of that data is now no longer accessible.

@@

\begin{figure}[!b]
 \includegraphics[width=3in]{img/clutterbuck.png} 
 \caption{Cost of change from an agile case study. From~\cite{Clutterbuck09}.}\label{fig:clutterbuck}
 \end{figure}
